Setting up Data

### 2020-06-11 doc/MS_v2.Rmd
tracked down difference due to M1 p values

---
title: 'Sex-Specific Evolution of the Meiotic Recombination Rate'
date: '`r Sys.Date()`'
author: "April L. Peterson, Bret Payseur"
output:
  word_document:
  pdf_document:
    toc: yes
    toc_depth: '5'
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
always_allow_html: yes
bibliography: refs_Main_Report.bib
---

@@

We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < `r 1/10000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p= `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count. After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex. A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = 3.9910^{-6} -- p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ). In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).

We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4). In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[8]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][10]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).

These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains. Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[2,][4]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M2$coefficients[5,][4]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[3,][4]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M2$coefficients[4,][4]` ). Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males. Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.

>>>>

---
title: 'Sex-Specific Evolution of the Meiotic Recombination Rate'
date: '`r Sys.Date()`'
author: "April L. Peterson, Bret Payseur"
output:
  html_document:
    toc: yes
    toc_depth: '5'
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '5'
  word_document:
always_allow_html: yes
bibliography: refs_Main_Report.bib
---

@@

We began with a full mixed model (M1; see Materials and Methods), showing that strain (random effect p < `r 1/10000`), sex (p = `r MLH1.MMlmer_results$sex.results`), subspecies (p= `r MLH1.MMlmer_results$sub.sp`), and subspecies*sex (p = `r MLH1.MMlmer_results$interaction.pval` ) each significantly affect MLH1 focus count. After a model including all factors as fixed effects (M2) revealed only weak contributions of subspecies, we focused on additional models designed to illuminate the role of strain and sex. A model with only these two variables (M3) identified two strains with particularly strong effects on MLH1 focus count:  _musculus^MSM^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][6]`) and _domesticus^G^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][3]` ). In addition, two strains exhibit strain-by-sex interactions: _molossinus^MSM^_  (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][13]` ) and _musculus^PWD^_ (p = `r summ.MLH1.M3.strain.fixed$coefficients[,4][12]` ).

We next fit models separately for  `r length(male_mouse.avs_for.analysis$mouse)` males and `r length(female_mouse.avs_for.analysis$mouse)` females (M4). In the male dataset, three strains affect MLH1 focus count (as observed in Figure 1C): _musculus^PWD^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][7]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[7]` foci), and _musculus^SKIVE^_ (p = `r sum_MLH1.male.glm.M4$coefficients[,4][9]` ; effect = `r sum_MLH1.male.glm.M4$coefficients[10]` ), and _molossinus^MSM^_  (p = `r sum_MLH1.male.glm.M4$coefficients[,4][8]`; effect = `r sum_MLH1.male.glm.M4$coefficients[9]`).

These three strains point to substantial evolution in the genome-wide recombination rate in spermatocytes; we subsequently refer to them as “high-recombination” strains. Analysis of the female dataset identifies four strains with effects on recombination rate: _domesticus^G^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][2]`), _molossinus^MSM^_  (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][5]`), _domesticus^LEW^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][3]` ), and _musculus^PWD^_ (p = `r sum_MLH1.female.spc.glm.M5$coefficients[,4][4]` ). Strain effect sizes in females are modest in magnitude (ranging from 1 to 4 foci) compared to those in males. Together, these results demonstrate that the genome-wide recombination rate evolves in a highly sex-specific manner.

### 2020-06-11 src/Result_Setup.Rmd
tracked down difference due to M1 p values

#DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]
#DF.HetC.MixedModel

@@

#HMI
HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
#unique(HetC.HMI$mouse)
#non-passing mice
#X.remove.mice <- c("", "")
#HetC.SPRET <- HetC.SPRET[ ! HetC.SPRET$mouse %in% SPRET.remove.mice, ]
#calq sex mean
HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)

@@

#PERC
HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]
#unique(HetC.PERC$mouse)
#non-passing mice
#G.remove.mice <- c("8jun15_G_f3","8jun15_G_f4")
#HetC.G <- HetC.PERC[ ! HetC.PERC$mouse %in% PERC.remove.mice, ]

@@

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

#calq sex mean
CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)

CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]

@@

#non-passing mice
#PWD.remove.mice <- c("1apr15_PWD_f2","8oct14_PWD_f2","8oct14_PWD_f5")
#HetC.PWD <- HetC.PWD[ ! HetC.PWD$mouse %in% PWD.remove.mice, ]

@@

#MOLF
HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]
unique(HetC.MOLF$mouse)

@@

#calq sex mean
MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)

MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]

```{r MLH1.HetC.plot, echo=FALSE, warning=FALSE}
#HETC plot
#change the main data?

HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length( nMLH1.foci),
                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd( nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#
#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]

HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length( nMLH1.foci),
                         mean_co = format(round(  mean( nMLH1.foci), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci),3), nsmall=3),
                         sd   = round(sd( nMLH1.foci), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]

#remove the 'other'

HetC_strain_table.M$sex <- "male"
HetC_strain_table.F$sex <- "female"

HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")

HetC_strain_table <- add_subsp(HetC_strain_table)

#HetC_strain_table$subsp <- c("Dom", "Dom", "Dom", "Musc", "Musc","Musc","Musc", "Cast", "Outgroup","Outgroup")

HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
#str(HetC_strain_table)

@@

                        Nmice = length(Nmice),
                        Ncells  = sum(Ncells), #might not to be sum
                        mean.SC = mean(mean_SC),
                        var.SC = mean(var), #mouse average variance of bivalent per mouse

                        #chrm classes
                        av.proportion_0CO = mean(proportion_0CO),
                        av.proportion_1CO = mean(proportion_1CO),
                        av.proportion_2CO = mean(proportion_2CO),
                        av.proportion_3CO = mean(proportion_3CO),

                        #1CO pos
                        mean.1CO.pos = mean(mean_1CO.pos),

                        #mean.IFD
                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS),
                        mean.IFD.PER = mean(mean_IFD.2CO_PER)
                        )

#add
strain.BivData.Table <- add_subsp(strain.BivData.Table)
strain.BivData.Table <- add_species(strain.BivData.Table)
```

@@

#showing the table..
#data.frame(table( droplevels( Curated_BivData  %>%  filter(hand.foci.count  == 0 | hand.foci.count  == 1 |                                       hand.foci.count  == 2 | hand.foci.count  == 3 | hand.foci.count  == 4 ) )$category ) )
```

RENAME the curate bivdata below

```{r Q1.biv.data_table.4MM.Male, echo=FALSE, eval=FALSE}

#this is redundant

male.av_4MM <- ddply(cleanCurated_BivData, c("mouse"), summarise,
                        Nmice = length(unique(mouse)),
                        Ncells  = length(Obj.ID),

                      # mean_CO = mean()

                        mean_SC = as.numeric(format(round(  mean(chromosomeLength), 3 ), nsmall=3) ),
                        cV = cv(chromosomeLength),
                        var = as.numeric(format(round(   var(chromosomeLength),3), nsmall=3) ),
                        sd   = round(sd(chromosomeLength), 3),
                        se   = round(sd / sqrt(Ncells), 3),

                       #calq mean.sc by chrm class
                       mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                       mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                       mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                       mean.SC_4CO = mean(chromosomeLength[which(hand.foci.count == 4)], na.rm = TRUE),

                # IFD things
                #  mean_IFD_ABS = as.numeric(format(round(  mean(IFD1_ABS, na.rm = TRUE), 3 ), nsmall=3) ),
                # mean_IFD_PER = as.numeric(format(round(  mean(IFD1_PER, na.rm = TRUE), 3 ), nsmall=3) ),

                       #CO position things, requires the melt?
              norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)]),

              #sis.co.ten!!
              mean.siscoten = mean(SisCoTen, na.rm=TRUE),

              mean.siscoten_1CO = mean(SisCoTen[which(hand.foci.count == 1)], na.rm = TRUE),
              mean.siscoten_2CO = mean(SisCoTen[which(hand.foci.count == 2)], na.rm = TRUE),
              mean.siscoten_3CO = mean(SisCoTen[which(hand.foci.count == 3)], na.rm = TRUE),

              mean.telo.dist = mean(telo_dist, na.rm=TRUE),
              mean.cent.dist = mean(dis.cent, na.rm=TRUE)
)

#vari
male.av_4MM <- add_sex(male.av_4MM)
male.av_4MM <- add_strain(male.av_4MM)
male.av_4MM <- add_subsp(male.av_4MM)
male.av_4MM <- add_category(male.av_4MM)

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)

#REMOVE JUVI mice
juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
#mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]

@@

@@

@@

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data=DF.HetC.MixedModel.HQ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

#summary(MLH1.MM.M1)$coefficients

@@

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)

@@

Curated_BivData <- Curated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")

@@

Curated_BivData$subsp <- factor(Curated_BivData$subsp , ordered = F)
Curated_BivData$strain <- factor(Curated_BivData$strain , ordered = F)

chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= ( Curated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") ) )

@@

```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
#REAL.short_mouse_table
#REAL.long_mouse_table
#mouse.av.long.biv.table

Q1.av.long.biv_scatr_box   <- ggplot(REAL.long_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")

Q1.av.short.biv_scatr_box   <- ggplot(REAL.short_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
```

>>>>

@@

#HMI
HetC.HMI <- MLH1_data[(MLH1_data$strain == "HMI")&(MLH1_data$quality < 5),]
HMI.mean.MLH1.obs <- ddply(.data=HetC.HMI, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)

@@

#PERC
HetC.PERC <- MLH1_data[(MLH1_data$strain == "PERC")&(MLH1_data$quality < 5),]

@@

#calq sex mean
CZECH.mean.MLH1.obs <- ddply(.data=HetC.CZECH, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)
CZECH.HETC.val <- CZECH.mean.MLH1.obs$mean.MLH1[1] / CZECH.mean.MLH1.obs$mean.MLH1[2]

@@

@@

#MOLF
HetC.MOLF <- MLH1_data[(MLH1_data$strain == "MOLF")&(MLH1_data$quality < 5),]

@@

#calq sex mean
MOLF.mean.MLH1.obs <- ddply(.data=HetC.MOLF, 
                 .(category),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                 mean.MLH1 = mean( nMLH1.foci),
                 var.MLH1 = var( nMLH1.foci)
)
MOLF.HETC.val <- MOLF.mean.MLH1.obs$mean.MLH1[1] / MOLF.mean.MLH1.obs$mean.MLH1[2]
```

```{r DF.MLH1.HetC.plot, echo=FALSE, warning=FALSE, eval=FALSE}

#I don't think this block is needed.

#HETC plot
#change the main data?

HetC_strain_table.M <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "male"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length( nMLH1.foci),
                         mean_co = format(round(  mean( nMLH1.foci, na.rm = TRUE), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3),
                         sd   = round(sd( nMLH1.foci, na.rm = TRUE), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#
#HetC_strain_table.M <- HetC_strain_table.M[ -c(7, 4, 10,11,12, 14, 17, 18,19), ]

HetC_strain_table.F <- ddply(MLH1_data[(MLH1_data$quality < 5)&(MLH1_data$sex == "female"),], c("strain"), summarise,
                         Nmice = length(unique(mouse)),
                         Ncells  = length( nMLH1.foci),
                         mean_co = format(round(  mean( nMLH1.foci, na.rm = TRUE), 3 ), nsmall=3),
                         var = format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3),
                         sd   = round(sd( nMLH1.foci, na.rm = TRUE), 3),
                         se   = round(sd / sqrt(Ncells), 3),
                         cV = round( (as.numeric(sd) / as.numeric(mean_co) ),3)
)
#HetC_strain_table.F <- HetC_strain_table.F[ -c(11), ]
#remove the 'other'
HetC_strain_table.M$sex <- "male"
HetC_strain_table.F$sex <- "female"

HetC_strain_table <- merge(HetC_strain_table.M, HetC_strain_table.F, by= "strain")
HetC_strain_table <- add_subsp(HetC_strain_table)

HetC_strain_table$mean_co.x <- as.numeric(HetC_strain_table$mean_co.x)
HetC_strain_table$mean_co.y <- as.numeric(HetC_strain_table$mean_co.y)
#str(HetC_strain_table)
```

@@

                       # Nmice = length(Nmice, na.rm = TRUE),
                        Ncells  = sum(Ncells, na.rm = TRUE), #might not to be sum
                        mean.SC = mean(mean_SC, na.rm =TRUE),
                        var.SC = mean(var, na.rm = TRUE), #mouse average variance of bivalent per mouse
                        #chrm classes
                        av.proportion_0CO = mean(proportion_0CO, na.rm = TRUE),
                        av.proportion_1CO = mean(proportion_1CO, na.rm = TRUE),
                        av.proportion_2CO = mean(proportion_2CO, na.rm = TRUE),
                        av.proportion_3CO = mean(proportion_3CO, na.rm = TRUE),
                        #1CO pos
                        mean.1CO.pos = mean(mean_1CO.pos, na.rm = TRUE),

                        #mean.IFD
                        mean.IFD.ABS = mean(mean_IFD.2CO_ABS, na.rm = TRUE),
                        mean.IFD.PER = mean(mean_IFD.2CO_PER, na.rm = TRUE)
                        )

strain.BivData.Table <- add_subsp(strain.BivData.Table)
strain.BivData.Table <- add_species(strain.BivData.Table)
#not sure the source of the NAs
```

@@

@@

#make sure that MOLF female is included!

#DF.HetC.MixedModel <- add_species(DF.HetC.MixedModel)
DF.HetC.MixedModel <- Mouse.Table_HQ[Mouse.Table_HQ$species == "M.musculus",]

DF.HetC.MixedModel.3 <- Mouse.Table_HQ %>% filter(species == "M.musculus")
#DF.HetC.MixedModel

127 mice, 8 strains (correct), 0 MOLF females!

comes from Mouse.Table_HQ, this has 0 females,

@@

there is another set of of this DF at line 200, this DF isn't defined in the setup script

DF.HetC.MixedModel.HQ should have only the relevant strains, but adding in

@@

MLH1.MM.M1 <- lmer(mean_co ~ subsp * sex + (1|strain), data= DF.HetC.MixedModel.HQ ) 
MLH1.MM.M1.reduce.sex <- lmer( mean_co ~ subsp  +(1|strain), data=DF.HetC.MixedModel.HQ )
MLH1.MM.M1.reduce.subp <- lmer( mean_co ~ sex  +(1|strain), data=DF.HetC.MixedModel.HQ )

@@

MLH1.MMlmer_results <- data.frame(sub.sp = MLH1.MM_subp, 
sex.results =  MLH1.MM_sex, 
               interaction.pval = MLH1.MM_interaction,  rand.pval = 
MLH1.MM_strain)
#I get the better values: when there is MOLF female and the old mice are removed
#consider removing these from the start
MLH1.MMlmer_results
```

**these should be the M1 MLH1 results**
sub.sp      sex.results      interaction.pval     rand.pval
1 0.0011     4.9e-06          0.00021         0

@@

cleanCurated_BivData <- cleanCurated_BivData %>% filter(Obj.ID != "13sep18_28jun18_MOLF_m1_sp1_18.1_rev_14")

@@

cleanCurated_BivData$subsp <- factor(cleanCurated_BivData$subsp , ordered = F)
cleanCurated_BivData$strain <- factor(cleanCurated_BivData$strain , ordered = F)

chrm.1CO.pos.chrmLength <- glm(PER_Foci_1 ~ chromosomeLength * subsp * strain, data= ( cleanCurated_BivData %>% filter(sex == "male")  %>%  filter(hand.foci.count == "1") ) )

@@

```{r Q1.mouse.av.F1.pos.LONG.BIV_scatter, echo=FALSE, eval=FALSE}
#REAL.short_mouse_table -- Short.biv_mouse_table
#REAL.long_mouse_table -- Long.biv_mouse_table
#mouse.av.long.biv.table

Q1.av.long.biv_scatr_box   <- ggplot(Long.biv_mouse_table  , aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nLong Biv")

Q1.av.short.biv_scatr_box   <- ggplot(Short.biv_mouse_table, aes(x=sex, y=mean_F1pos_PER, color=strain) )+geom_jitter( )+ geom_boxplot(aes(y= mean_F1pos_PER ),alpha=.3 ) + scale_color_manual(values=colors_of_strains)+facet_wrap(~subsp, scales = "free")+ylim(c(0,1.1))+ggtitle("Mouse av. \nNormalized 1CO Foci pos \nShort Biv")
```

### 2020-06-11 src/SetupData.Rmd
tracked down difference due to M1 p values

@@

#juvi.mice_DF <- AP_mouse_table_w.Ages[(AP_mouse_table_w.Ages$age.days < 15) & (AP_mouse_table_w.Ages$sex.y == "male"),]
#juvi.mice_DF <- juvi.mice_DF[!(is.na(juvi.mice_DF$mouse) | juvi.mice_DF$mouse==""), ]

@@

#change name to Mouse.Table_HQ from #AP_mouse_table.MLH1_HQ
#

@@

#remove juvi mice via the list
#remove bad ages
#this is removing female MOLF?

#this removes everything but males in these categories
#Mouse.Table_HQ <- Mouse.Table_HQ[ (Mouse.Table_HQ$sex == "female") | ( #(Mouse.Table_HQ$sex == "male") & (Mouse.Table_HQ$age.weeks < 25) & #(Mouse.Table_HQ$age.days != 13) ), ]

#i don't understand why so many WSB female are removed..
# | is.na(Mouse.Table_HQ$age.weeks) | (Mouse.Table_HQ$age.weeks == 0)

@@

Mouse.Table_HQ <- Mouse.Table_HQ.OG

#during merging -- try to just merge a few cols: 
Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)
#this removes ~28 mice

@@

Mouse.Table_HQ <- Mouse.Table_HQ[ ! Mouse.Table_HQ$mouse %in% old.mice.df, ]

@@

```{r Age.Hists, echo=FALSE, warning=FALSE}
#age.weeks col is missing

>>>>

#remove all X's from MLH1
#add date imaged

@@

@@

@@

#there are ~26 old mice

#this doesn't remove
Mouse.Table_HQ <- Mouse.Table_HQ.OG[ ! Mouse.Table_HQ.OG$mouse %in% old.mice.list, ]

#MLH1_data <- MLH1_data[ ! MLH1_data$mouse %in% juvi_mice.list, ]

@@

#Mouse.Table_HQ.OG <- Mouse.Table_HQ

#during merging -- try to just merge a few cols: 
#don't merge
#Mouse.Table_HQ <- merge(Mouse.Table_HQ, meta.data)

@@

@@

```{r Age.Hists, echo=FALSE, warning=FALSE, eval=FALSE}
#age. weeks was removed from the Mouse.Table.HQ for now

#age.weeks col is missing
